Expert-In-The-Loop Causal Discovery: Iterative Model Refinement Using Expert Knowledge

### 2025-01-25 paper/main.tex
Udpates
Updates figure 4
Updates

There has been work on trying to come up with a single test to combine all
these test to give a single answer \citep{Eulig2023, Shipley2000}. However,
what is missing is the guide on how to how to fix these tests. To tackle these
issues with CI based testing, we take inspiration from \emph{modification
indices} in specification search. The modification indices method tries to
identify and rank changes in the model such that modifications that improve the
model fit the most are ranked higher.This ranking allows researchers to focus
first on the suggestions that can improve the model the most and while
integrating their expertise to choose the most optimal modifications. Analogous
to this, in this paper we develop a measure for partial association for mixed
data that allows us to rank the CI violations. Additionally, this measure of
conditional association also allows us tackle other issues with CI based
testing outlined in the previous paragraphs.

@@

\begin{figure}
\begin{subfigure}{0.5 \textwidth}
	\includegraphics[page=1]{figures.pdf}
\end{subfigure}
\begin{subfigure}{0.5 \textwidth}
	\includegraphics[page=2]{figures.pdf}
\end{subfigure}
\begin{subfigure}{0.5\textwidth}
	\includegraphics[page=3]{figures.pdf}
\end{subfigure}

@@

\begin{figure}
	\centering
	\includegraphics[scale=0.4]{plots/web_tool_full.png}
	\caption{A screenshot of the web tool for constructing the model. Users
		can upload their dataset after which the tool creates an empty
		graph and shows all pair of variables which are associated in
		the model using undirected red edges with the strength of
		association represented using edge width. Users can then add
		edges to the model (shown in green). Unnecessary edges are shown in 
		black.}
\end{figure}

@@

\begin{figure}
	\begin{subfigure}{0.5\textwidth}
		\centering
		\includegraphics{../code/fig3/shd_ribbon.pdf}
		\caption{}
	\end{subfigure}
	\begin{subfigure}{0.5\textwidth}
		\centering
		\includegraphics{../code/fig3/sid_ribbon.pdf}
		\caption{}
	\end{subfigure}
	\caption{Comparison of PC and Hill-Climb Search algorithms against
		manually drawn DAGs using the assistance method. As PC and
		Hill-Climb Search return the Markov Equivalence Class (MEC), we
		use the best and worst scoring orientation of the MEC to get
		the range of values. The human drawn values are done for
		multiple accuracy values of the oracle.}
\end{figure}

\begin{figure}
	\begin{subfigure}{0.5\textwidth}
		\centering
		\includegraphics{../code/fig4/unexplained_effect.pdf}
		\caption{}
	\end{subfigure}
	\begin{subfigure}{0.5\textwidth}
		\centering
		\includegraphics{../code/fig4/ll.pdf}
		\caption{}
	\end{subfigure}
	\caption{Plots showing improvements in fit of the model over iterations
	on the Adult income dataset. For the expert simulation an LLM is used
	for the edge orientations. (a) Shows using total unexplained effect that is the
	sum of effect size between variables that don't have an edge between them. (b)
	Shows the log-likelihood of the overall model.
	}
\end{figure}

@@

\begin{figure}
	\centering
	\begin{subfigure}{0.5\textwidth}
		\caption{\todo[inline]{Add the prompt used}}
	\end{subfigure}
	\begin{subfigure}{0.5\textwidth}
		\centering
		\includegraphics[scale=0.3]{../code/fig5/llm_adult.png}
		\caption{\todo[inline]{Change figure to have consistent style}}
	\end{subfigure}
\end{figure}

>>>>

There has been work on trying to come up with a single test to combine all
these test to give a single answer \citep{Eulig2023, Shipley2000}. However,
what is missing is the guide on how to how to fix these tests. Additionally
there has been a lot of work to integrate expert knowledge into causal
discovery algorithms which has shown significant improvement. To tackle these
issues with CI based testing, we take inspiration from \emph{modification
indices} in specification search. The modification indices method tries to
identify and rank changes in the model such that modifications that improve the
model fit the most are ranked higher.This ranking allows researchers to focus
first on the suggestions that can improve the model the most and while
integrating their expertise to choose the most optimal modifications. Analogous
to this, in this paper we develop a measure for partial association for mixed
data that allows us to rank the CI violations. Additionally, this measure of
conditional association also allows us tackle other issues with CI based
testing outlined in the previous paragraphs.

@@

\begin{figure}
\begin{subfigure}{0.25 \textwidth}
	\includegraphics[page=1]{figures.pdf}
\end{subfigure}%
\begin{subfigure}{0.25 \textwidth}
	\includegraphics[page=2]{figures.pdf}
\end{subfigure}
\begin{subfigure}{0.25 \textwidth}
	\includegraphics[page=3]{figures.pdf}
\end{subfigure}%
\begin{subfigure}{0.25\textwidth}
	\includegraphics[page=4]{figures.pdf}
\end{subfigure}

@@

\begin{figure}
	\centering
	\includegraphics[scale=0.4]{../code/plots/web_tool_full.png}
	\caption{A screenshot of the web tool for constructing the model. Users
		can upload their dataset after which the tool creates an empty
		graph and shows all pair of variables which are associated in
		the model using undirected red edges with the strength of
		association represented using edge width. Users can then add
		edges to the model (shown in green). Unnecessary edges are shown in 
		black.}
\end{figure}

@@

\begin{figure}
	\centering
	\begin{subfigure}{0.5\textwidth}
		\centering
		\includegraphics{../code/fig3/shd_ribbon.pdf}
		\caption{}
	\end{subfigure}
	\begin{subfigure}{0.5\textwidth}
		\centering
		\includegraphics{../code/fig3/sid_ribbon.pdf}
		\caption{}
	\end{subfigure}
	\caption{Comparison of PC and Hill-Climb Search algorithms against
		manually drawn DAGs using the assistance method. As PC and
		Hill-Climb Search return the Markov Equivalence Class (MEC), we
		use the best and worst scoring orientation of the MEC to get
		the range of values. The human drawn values are done for
		multiple accuracy values of the oracle.}
\end{figure}

\begin{figure}
	\begin{subfigure}{0.25\textwidth}
		\centering
		\includegraphics{../code/fig4/unexplained_effect.pdf}
		\caption{}
	\end{subfigure}%
	\begin{subfigure}{0.25\textwidth}
		\centering
		\includegraphics{../code/fig4/ll.pdf}
		\caption{}
	\end{subfigure}
	\caption{Plots showing improvements in fit of the model over iterations
	on the Adult income dataset. For the expert simulation an LLM is used
	for the edge orientations. (a) Shows using total unexplained effect that is the
	sum of effect size between variables that don't have an edge between them. (b)
	Shows the log-likelihood of the overall model.
	}
\end{figure}

@@

\begin{figure}
	\centering
	\begin{subfigure}{0.5\textwidth}
		\caption{\todo[inline]{Add the prompt used}}
	\end{subfigure}
	\begin{subfigure}{0.5\textwidth}
		\centering
		\includegraphics[scale=0.3]{../code/plots/llm_adult.png}
		\caption{\todo[inline]{Change figure to have consistent style}}
	\end{subfigure}
\end{figure}
